### 2018-06-25 VV570/HW5/submission.tex
basically done

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[paper=a4, fontsize=11pt]{scrartcl} % A4 paper and 11pt font size

\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\usepackage{fourier} % Use the Adobe Utopia font for the document - comment this line to return to the LaTeX default
\usepackage[english]{babel} % English language/hyphenation
\usepackage{amsmath,amsfonts,amsthm,amssymb} % Math packages
\usepackage{mathrsfs}

\usepackage{algorithm, algorithmic}
\renewcommand{\algorithmicrequire}{\textbf{Input:}} %Use Input in the format of Algorithm  
\renewcommand{\algorithmicensure}{\textbf{Output:}} %UseOutput in the format of Algorithm

\usepackage{listings}
\lstset{language=Matlab}

\usepackage{lipsum} % Used for inserting dummy 'Lorem ipsum' text into the template

\usepackage{sectsty} % Allows customizing section commands
\allsectionsfont{\centering \normalfont\scshape} % Make all sections centered, the default font and small caps

\usepackage{fancyhdr} % Custom headers and footers
\pagestyle{fancyplain} % Makes all pages in the document conform to the custom headers and footers
\fancyhead{} % No page header - if you want one, create it in the same way as the footers below
\fancyfoot[L]{} % Empty left footer
\fancyfoot[C]{} % Empty center footer
\fancyfoot[R]{\thepage} % Page numbering for right footer
\renewcommand{\headrulewidth}{0pt} % Remove header underlines
\renewcommand{\footrulewidth}{0pt} % Remove footer underlines
\setlength{\headheight}{13.6pt} % Customize the height of the header

\numberwithin{equation}{section} % Number equations within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{figure}{section} % Number figures within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{table}{section} % Number tables within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)

\setlength\parindent{0pt} % Removes all indentation from paragraphs - comment this line for an assignment with lots of text

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\newcommand{\horrule}[1]{\rule{\linewidth}{#1}} % Create horizontal rule command with 1 argument of height

\title{	
\normalfont \normalsize 
\textsc{Shanghai Jiao Tong University, UM-SJTU JOINT INSTITUTE} \\ [25pt] % Your university, school and/or department name(s)
\horrule{0.5pt} \\[0.4cm] % Thin top horizontal rule
\huge Introduction to Numerical Analysis \\ HW5 \\ % The assignment title
\horrule{2pt} \\[0.5cm] % Thick bottom horizontal rule
}

\author{Yu Cang \\ 018370210001} % Your name

\date{\normalsize \today} % Today's date or a custom date

\begin{document}

\maketitle % Print the title

%----------------------------------------------------------------------------------------
%	PROBLEM 1
%----------------------------------------------------------------------------------------
\section{Lebesgue constant for Chebyshev nodes}
\begin{enumerate}
	\item 
		\begin{enumerate}
			\item 
				\begin{proof}
					Denote
					\begin{equation}
						\begin{aligned}
							LHS &\triangleq (x-x_i)l_i(x) \\
							RHS &\triangleq \frac{T_{n+1}(x)}{T_{n+1}'(x_i)}
						\end{aligned}
					\end{equation}
					It is left to prove $LHS = RHS$.\\
					The left part can be written as
					\begin{equation}
						LHS = c_l \omega(x)
					\end{equation}
					where 
					\begin{equation}
						\omega(x) = \prod_{i=0}^{n}(x-x_i)
					\end{equation}
					and
					\begin{equation}
						c_l = \frac{1}{\prod_{k=0, k\neq i}^{n}(x_i - x_k)} 
					\end{equation}

					Since both $LHS$ and $RHS$ are polynomials of order $n+1$, they are equivalent iff. they have same roots and leading coefficients.\\
					On the one hand,as for all $x_i$, where $i = 0, 1, ... , n$
					\begin{equation}
						T_{n+1}(x_i) = cos((n+1)\theta_i) = cos(\frac{2i+1}{2}\pi) = 0
					\end{equation}
					Thus, $LHS$ and $RHS$ have same roots. $RHS$ can therefore be written as
					\begin{equation}
						RHS(x) = c_r \omega(x)
					\end{equation}
					On the other hand, since
					\begin{equation}
						LHS'(x)|_{x=x_i} = (l_i(x) + (x-x_i)l_i'(x))|_{x=x_i} = 1
					\end{equation}
					and
					\begin{equation}
						RHS'(x)|_{x=x_i} = \frac{T'_{n+1}(x)}{T_{n+1}'(x_i)}\Big|_{x=x_i} = 1
					\end{equation}
					Thus, the leading coefficients of $LHS$ and $RHS$ are equal, namely $c_l = c_r$.\\
					Hence, $LHS=RHS$.
				\end{proof}
			\item 
				\begin{proof}
					\begin{equation}
						\begin{aligned}
							T'_{n+1}(x) & = (cos((n+1)arccos(x)))'\\
							            & = sin((n+1)arccos(x)) (n+1) \frac{1}{\sqrt{1-x^2}}\\
							            & = \frac{n+1}{\sqrt{1-cos^2(\theta)}} sin((n+1)\theta) 
						\end{aligned}
					\end{equation}
					As $\theta_k = \frac{2k+1}{2(n+1)}\pi$, thus, $sin(\theta_k) > 0$, and
					\begin{equation}
						T'_{n+1}(x_k) = \frac{n+1}{sin(\theta_k)} sin(\frac{2k+1}{2}\pi) = {(-1)}^k \frac{n+1}{sin(\theta_k)}
					\end{equation}
				\end{proof}
			\item
				\begin{proof}
					As
					\begin{equation}
						T_{n+1}(1) = cos((n+1)arccos(1)) = 1	
					\end{equation}
					Thus
					\begin{equation}
						\begin{aligned}
							\sum_{i=0}^{n}|l_i(1)| & = \sum_{i=0}^{n}\Bigg|\frac{T_{n+1}(1)}{(1-x_i)T'_{n+1}(x_i)}\Bigg| \\
												   & = \sum_{i=0}^{n} \frac{1}{\Big|(1-x_i)T'_{n+1}(x_i)\Big|}\\
												   & = \frac{1}{n+1} \sum_{i=0}^{n} \bigg|\frac{sin\theta_i}{(1-cos\theta_i)} \bigg|\\
												   & = \frac{1}{n+1} \sum_{i=0}^{n} \bigg|\frac{sin\theta_i}{2sin^2(\frac{\theta_i}{2})} \bigg|\\
												   & \geq  \frac{1}{n+1} \sum_{i=0}^{n} cot(\frac{\theta_i}{2})\\
						\end{aligned}
					\end{equation}
				\end{proof} 			
		\end{enumerate}
	\item 
		\begin{enumerate}
			\item 
				\begin{proof}
					According to the mean value theorem, there exists $\theta \in [\frac{\theta_k}{2}, \frac{\theta_{k+1}}{2}]$, s.t.
					\begin{equation}
						\int_{\frac{\theta_k}{2}}^{\frac{\theta_{k+1}}{2}} cot(t) dt = \frac{\theta_{k+1} - \theta_k}{2} cot(\theta)
					\end{equation}
					As $cot'(t) = -\frac{1}{sin^2(t)} < 0$, and $\theta_k \leq \theta \leq \theta_{k+1}$, thus
					\begin{equation}
						cot(\theta) \leq cot(\theta_k)
					\end{equation}
					Therefore
					\begin{equation}
						\int_{\frac{\theta_k}{2}}^{\frac{\theta_{k+1}}{2}} cot(t) dt \leq \frac{\theta_{k+1} - \theta_k}{2} cot(\theta_k)
					\end{equation}
				\end{proof}

			\item 
				\begin{proof}
					As $\theta_{k+1} - \theta_k = \frac{\pi}{n+1}$ and according to the result that have been proved above
					\begin{equation}
						\begin{aligned}
							\sum_{k=0}^{n} \int_{\frac{\theta_k}{2}}^{\frac{\theta_{k+1}}{2}} cot(t) dt 
							& \leq \sum_{k=0}^{n} \frac{\theta_{k+1} - \theta_k}{2}cot(\frac{\theta_k}{2})\\
							& = \frac{\pi}{2(n+1)} \sum_{k=0}^{n} cot(\frac{\theta_k}{2})
						\end{aligned}
					\end{equation}
				\end{proof}

			\item 
				\begin{proof}
					As $\theta_n = \frac{2n+1}{2n+2}\pi < \pi$, $\theta_{n+1} = \frac{2n+3}{2n+2} > \pi$, and $cot(x)$ is positive over $(0, \frac{\pi}{2})$, while negative otherwise.
					Thus
					\begin{equation}
							\int_{\frac{\theta_0}{2}}^{\frac{\pi}{2}} cot(t) dt 
							\leq \int_{\frac{\theta_0}{2}}^{\frac{\theta_n}{2}} cot(t) dt
							= \sum_{k=0}^{n-1} \int_{\frac{\theta_k}{2}}^{\frac{\theta_{k+1}}{2}} cot(t) dt
					\end{equation}
					Hence
					\begin{equation}
						\int_{\frac{\theta_0}{2}}^{\frac{\pi}{2}} cot(t) dt \leq \frac{\pi}{2(n+1)} \sum_{i=0}^{n}cot(\frac{\theta_i}{2})
					\end{equation}
					(... not fine)
				\end{proof}

		\end{enumerate}

	\item 
		\begin{proof}
			\begin{equation}
				\begin{aligned}
					\Lambda_n & = \max_{x \in [a, b]} \sum_{i=0}^{n} |l_i(x)|\\
							  & \geq \sum_{i=0}^{n} |l_i(1)|\\
							  & \geq \frac{1}{n+1} \sum_{i=0}^{n} cot(\frac{\theta_i}{2})\\
							  & \geq \frac{2}{\pi} \int_{\theta_0 / 2}^{\pi/2} cot(t)dt\\
							  & = \frac{2}{\pi} ln(|sin(t)|) \Bigg| _{\theta_0/2}^{\pi/2}\\
							  & = - \frac{2}{\pi} ln(sin(\frac{\theta_0}{2}))\\
							  & \geq \frac{2}{\pi} ln(\frac{2}{\theta_0}) = \frac{2}{\pi} ln(\frac{4(n+1)}{\pi})\\
							  & \ge \frac{2}{\pi} ln(n)
				\end{aligned}
			\end{equation}
		\end{proof}

\end{enumerate}

%----------------------------------------------------------------------------------------
%	PROBLEM 2
%----------------------------------------------------------------------------------------
\section{Interpolation}
\begin{enumerate}
	\item 
		\begin{proof}
			Since $\Phi(f)=0$, $f(x_k) + f(y_k) = 0$ is valid for any $k$.\\
			Thus $f(x_k) = f(y_k) = 0$ or $f(x_k)$ and $f(y_k)$ have different signs.\\
			In the first case, take $\xi_k = x_k$ or $\xi_k = y_k$;\\
			In the second case, as $f$ is continous over $[a, b]$, according to the  intermediate value theorem, there exists $\xi_k \in [x_k, y_k]$ s.t. $f(\xi_k) = 0$.
		\end{proof}

	\item 
		\begin{proof}
			Let $h_1(x) \in \mathbb{R}_n [x]$, $h_2(x) \in \mathbb{R}_n [x]$ and  $h_1(x) \neq h_2(x)$. They can be written as
			\begin{equation}
				h_1(x) = \sum_{i=0}^{n} a_i x^i
			\end{equation}
			\begin{equation}
				h_2(x) = \sum_{i=0}^{n} b_i x^i
			\end{equation}
			Assume $\Phi(h_1) = \Phi(h_2)$, then
			\begin{equation}
				h_1(x_k) + h_1(y_k) = h_2(x_k) + h_2(y_k) \ \ (k=0, 1, ... , n)
			\end{equation}
			Thus
			\begin{equation}
				 \begin{bmatrix}
					(1+1) & (x_0 + y_0) &  (x_0^2 + y_0^2) & ... & (x_0^n + y_0^n)\\
					(1+1) & (x_1 + y_1)  &  (x_1^2 + y_1^2) & ... & (x_1^n  + y_1^n)\\
					\vdots  & \vdots    & \vdots & \ddots & \vdots \\
					(1+1) & (x_n + y_n) & (x_n^2 + y_n^2) & ... & (x_n^n + y_n^n)\\
				\end{bmatrix}
				\begin{bmatrix}
					a_0 - b_0\\
					a_1 - b_1\\
					\vdots \\
					a_n - b_n\\
				\end{bmatrix}
				=
				\begin{bmatrix}
					0\\
					0\\
					\vdots \\
					0\\
				\end{bmatrix}
			\end{equation}
			The determinant of the coefficient matrix above is positive as it is the sum of $2^{n+1}$ determinents, each of these be positive as $x_0 < y_0 < x_1 < y_1 < ... < x_n < y_n$. Thus, solution to this linear system is $x = [0, 0, ... , 0]^T$, which means
			\begin{equation}
				a_i = b_i \ \ (i=0, 1, ... , n)
			\end{equation}
			So, the hypothsis fails as $h_1(x) = h_2(x)$. Hence, $\Phi(f) \neq \Phi(g)$, which means $\Phi$ is injective.\\
			Assume $\Phi(P1) = \Phi(f)$, $\Phi(P2) = \Phi(f)$, according to the converse-negative proposition of the injective property, there should be $P1 = P2$, thus the unicity is proved.
		\end{proof}

	\item 
		\begin{proof}
			Let $P_n(x) = a_n x^n + a_{n-1} x^{n-1} + ... + a_1 x + a_0$, denote $h(x) = P_n(x) - f(x)$. As $\Phi(P_n) = \Phi(f)$, then
			\begin{equation}
			h(x_k) + h(y_k) = 0
			\end{equation}
			Thus, there exists $\xi_k \in [x_k, y_k]$ s.t. $h(\xi_k) = 0$. So
			\begin{equation}
			\sum_{i=0}^{n}a_i \xi_k ^ i = f(\xi_k) \ \ (k=0, 1, ... , n)
			\end{equation}
			which can be written as matrix form
			\begin{equation}
			\begin{bmatrix}
				1 & \xi_0 & \xi_0^2 & ... & \xi_0^n \\
				1 & \xi_1 & \xi_1^2 & ... & \xi_1^n \\
				\vdots  & \vdots & \vdots & \ddots & \vdots \\
				1 & \xi_n & \xi_n^2 & ... & \xi_n^n \\
			\end{bmatrix}
			\begin{bmatrix}
				a_0\\
				a_1\\
				\vdots \\
				a_n\\
			\end{bmatrix}
			=
			\begin{bmatrix}
				f(\xi_0)\\
				f(\xi_1)\\
				\vdots \\
				f(\xi_n)\\
			\end{bmatrix}
			\end{equation}
			The determinant of the coefficient matrix is the so called Vandermonde determinant, and $\xi_j > \xi_i$ when $j>i$. Thus, determinant of the coefficient matrix is positive, which means there exist $n+1$ points in $[a, b]$, denoted as $\xi_i$, s.t. $P(\xi_i) = f(\xi_i)$. Thus $P_n$ is the interpolation polynomial of $f$.\\

			Applying the Taylor theorem, it's clear that 
			\begin{equation}
				f(x)-P_n(x) = \frac{(x-a)^{n+1}}{(n+1)!}f^{(n+1)}(\xi)
			\end{equation}
			Thus
			\begin{equation}
				\lVert f- p_n \rVert_\infty = \max_{x \in [a, b]} |f-P_n| \leq \frac{(b-a)^{n+1}}{(n+1)!}\sup_{x \in [a, b]}|f^{(n+1)}(x)|
			\end{equation}

		\end{proof}

\end{enumerate}

%----------------------------------------------------------------------------------------
%	PROBLEM 3
%----------------------------------------------------------------------------------------
\section{Trigonometric polynomials}
\begin{enumerate}
	\item 
		\begin{proof}
			It's clear that the statement holds when $k=0, 1$ as $1$ and $cos \theta$ are in $T_0$ and $T_1$ respectively.\\
			Suppose the statement is still valid when $k = n-1$, thus, there exists $Q_{n-1}$ s.t.
			\begin{equation}
				(cos\theta)^{n-1} = Q_{n-1}
			\end{equation}
			When it comes to $k = n$, with the induction hypothesis,
			\begin{equation}
				\begin{aligned}
						(cos\theta)^n & =  Q_{n-1} \times cos\theta\\
						                     & = (\frac{a_0}{\sqrt{2}} + \sum_{k=1}^{n-1}a_k cosk\theta) cos\theta\\
						                     & = \frac{a_0}{\sqrt{2}}cos\theta + \sum_{k=1}^{n-1}a_k (\frac{cos(k+1)\theta + cos(k-1)\theta}{2})\\
						                     & = \frac{b_0}{\sqrt{2}} + \sum_{k=0}^{n} b_k cosk\theta \triangleq Q_n
				\end{aligned}
			\end{equation}
			Thus, $(cos\theta)^n$ is in $T_n$. Hence, for any $0 \leq k \leq n$, $(cos\theta)^k$ is in $T_n$ as $T_k \subseteq T_n$.\\

			(the bijection part haven't figured out yet ...)

		\end{proof}
	\item 
		\begin{proof}
			As $Q_n(\theta_i) = F(\theta_i)$, the existence of $Q_n \in T_n$ is equivalent to the existence of solution of the linear system given as below
			\begin{equation}
				\begin{bmatrix}
					\frac{1}{\sqrt{2}} & cos(\theta_0) & cos(2\theta_0) & ... & cos(n\theta_0) \\
					\frac{1}{\sqrt{2}} & cos(\theta_1) & cos(2\theta_1) & ... & cos(n\theta_1) \\
					\vdots                 & \vdots           & \vdots             & \ddots  & \vdots     \\
					\frac{1}{\sqrt{2}} & cos(\theta_n) & cos(2\theta_n) & ... & cos(n\theta_n) \\
				\end{bmatrix}
				\begin{bmatrix}
					a_0 \\ a_1 \\ a_2 \\ \vdots \\ a_n\\
				\end{bmatrix}
				=
				\begin{bmatrix}
					F(\theta_0) \\ F(\theta_1) \\ F(\theta_2) \\ \vdots \\ F(\theta_n)\\
				\end{bmatrix}
			\end{equation}
			Thus, it's left to prove the determinant of the coefficient matrix is non-zero, which can be denoted as $\det(A)$.\\
			As $cos(k\theta)$ can be written as
			\begin{equation}
				cos(k\theta) = \sum_{n=0}^{k} b_n(cos\theta)^n
			\end{equation}
			Thus, $\det(A)$ can be simplified as below, where $C$ is an non-zero factor.
			\begin{equation}
				\det(A) = C \det
				\begin{bmatrix}
					1 & cos(\theta_0) & cos^2(\theta_0) & ... & cos^n(\theta_0) \\
					1 & cos(\theta_1) & cos^2(\theta_1) & ... & cos^n(\theta_1) \\
					\vdots                 & \vdots           & \vdots             & \ddots  & \vdots     \\
					1 & cos(\theta_n) & cos^2(\theta_n) & ... & cos^n(\theta_n) \\
				\end{bmatrix}
			\end{equation}
			With $\theta_i = \frac{2i+1}{2(n+1)}\pi$, it's clear that $\det(A)\neq0$ as its Vandermon derterminant part is non-zero. Thus the existence of $Q_n$ is valid.
		\end{proof}

	\item 
		\begin{proof}
			It has been exactly proved above.
		\end{proof}

	\item 
		\begin{proof}

		\end{proof}

\end{enumerate}

\end{document}