### 2018-09-23 VV556/HW2/submission.tex
add part of ex11

\documentclass[paper=a4, fontsize=11pt]{scrartcl} % A4 paper and 11pt font size

\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\usepackage{fourier} % Use the Adobe Utopia font for the document - comment this line to return to the LaTeX default
\usepackage[english]{babel} % English language/hyphenation
\usepackage{amsmath,amsfonts,amsthm,amssymb} % Math packages

\usepackage{algorithm, algorithmic}
\renewcommand{\algorithmicrequire}{\textbf{Input:}} %Use Input in the format of Algorithm  
\renewcommand{\algorithmicensure}{\textbf{Output:}} %UseOutput in the format of Algorithm

\usepackage{listings}
\lstset{language=Matlab}

\usepackage{lipsum} % Used for inserting dummy 'Lorem ipsum' text into the template

\usepackage{sectsty} % Allows customizing section commands
\allsectionsfont{\centering \normalfont\scshape} % Make all sections centered, the default font and small caps

\usepackage{fancyhdr} % Custom headers and footers
\pagestyle{fancyplain} % Makes all pages in the document conform to the custom headers and footers
\fancyhead{} % No page header - if you want one, create it in the same way as the footers below
\fancyfoot[L]{} % Empty left footer
\fancyfoot[C]{} % Empty center footer
\fancyfoot[R]{\thepage} % Page numbering for right footer
\renewcommand{\headrulewidth}{0pt} % Remove header underlines
\renewcommand{\footrulewidth}{0pt} % Remove footer underlines
\setlength{\headheight}{13.6pt} % Customize the height of the header

\numberwithin{equation}{section} % Number equations within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{figure}{section} % Number figures within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{table}{section} % Number tables within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)

\setlength\parindent{0pt} % Removes all indentation from paragraphs - comment this line for an assignment with lots of text

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\newcommand{\horrule}[1]{\rule{\linewidth}{#1}} % Create horizontal rule command with 1 argument of height

\title{	
\normalfont \normalsize 
\textsc{Shanghai Jiao Tong University, UM-SJTU JOINT INSTITUTE} \\ [25pt] % Your university, school and/or department name(s)
\horrule{0.5pt} \\[0.4cm] % Thin top horizontal rule
\huge Methods of Applied Mathematics I\\ HW2 \\ % The assignment title
\horrule{2pt} \\[0.5cm] % Thick bottom horizontal rule
}

\author{Yu Cang \\ 018370210001} % Your name

\date{\normalsize \today} % Today's date or a custom date

\begin{document}

\maketitle % Print the title

\section{Exercis2.1}
	\begin{enumerate}
		\item 
			\begin{proof}
				\begin{equation}
					\begin{aligned}
						RHS & = \frac{1}{4} (||x+y||^2  - ||x-y||^2)\\
						    & = \frac{1}{4} (<x+y, x+y> - <x-y, x-y>)\\
						    & = \frac{1}{4} (<x+y, x> + <x+y, y> - <x-y, x> - <x-y, -y>)\\
						    & = \frac{1}{4} (\overline{<x, x+y>} + \overline{<y, x+y>} - \overline{<x, x-y>} + \overline{<y, x-y>})\\
						    & = \frac{1}{4} (\overline{<x, x>} + \overline{<x, y>} + \overline{<y, x>} + \overline{<y, y>} - \overline{<x, x>} + \overline{<x, y>} + \overline{<y, x>} - \overline{<y, y>})\\
						    & = \frac{1}{2}(\overline{<x,y>} + \overline{<y, x>})\\
						    & = \frac{1}{2}(<x, y> + <y, x>)\\
						    & = <x, y> = LHS
					\end{aligned}
				\end{equation}
				The last line is valid as the inner-product is defined on real space s.t. $<x, y>=<y, x>$.
			\end{proof}
		\item 
			\begin{proof}
				As have been proved aboved
				\begin{equation}
					\begin{aligned}
						   & \frac{1}{4} (||x+y||^2  - ||x-y||^2)\\
						=  & \frac{1}{2}(<x, y> + <y, x>)
					\end{aligned}
				\end{equation}
				Also, replace $y$ with $iy$ yields
				\begin{equation}
					\begin{aligned}
						  & \frac{i}{4} (||x+iy||^2  - ||x-iy||^2)\\
						= & \frac{i}{2} (<x, iy> + <iy, x>)\\
						= & \frac{i}{2} (i<x, y> + \overline{i}<y, x>)\\
						= & \frac{1}{2} (-<x, y> + <y,x>)
					\end{aligned}
				\end{equation}
				Thus
				\begin{equation}
					\begin{aligned}
						RHS & = \frac{1}{4} (||x+y||^2  - ||x-y||^2) - \frac{i}{4} (||x+iy||^2  - ||x-iy||^2)\\
						    & = \frac{1}{2}(<x, y> + <y, x>) - \frac{1}{2} (-<x, y> + <y,x>)\\
						    & = <x, y> = LHS
					\end{aligned}
				\end{equation}
			\end{proof}
		\item 
			\begin{proof}
				\begin{equation}
					\begin{aligned}
						LHS & = ||x+y||^2 + ||x-y||^2\\
						    & = <x+y, x+y> + <x-y, x-y>\\
						    & = <x+y, x> + <x+y, y> + <x-y, x> - <x-y, y>\\
						    & = \overline{<x, x+y>} + \overline{<y, x+y>} + \overline{<x, x-y>} - \overline{<y, x-y>}\\
						    & = \overline{<x, x>} + \overline{<x, y>} + \overline{<y, x>} + \overline{<y, y>} + \overline{<x, x>} - \overline{<x, y>} - \overline{<y, x>} + \overline{<y, y>}\\
						    & = 2(<x, x>+<y, y>)\\
						    & = 2(||x||^2 + ||y||^2) = RHS
					\end{aligned}
				\end{equation}
			\end{proof}	
		\item 
			\begin{proof}

			\end{proof}

		\item

	\end{enumerate}

\section{Exercise2.2}
	\begin{enumerate}
		\item

	\end{enumerate}

\end{document}