### 2019-03-15 VG500/WEEK3/20190315.tex
weekly writting

\documentclass[paper=a4, fontsize=11pt]{scrartcl}

\usepackage{lipsum}
\usepackage{amsmath, amsfonts, amssymb}

\title{Summary of "Progress-variable approach for large-eddy simulation of non-premixed turbulent combustion"}
\author{Yu Cang \ \ 018370210001}
\date{2019-03-15}

\begin{document}

\maketitle

To simulate the combustion process, a set of partial differential equations needs to be solved, including the continuity equation, the momentum equation, the energy equation and the species equations. There exist quite strong non-linear source terms in the energy equation and the species equations, which pose difficulties to traditional solution methods.\newline
Instead of treating these non-linear terms directly, a new approach is introduced in this paper, where two mapping variables are introduced. In this way, all of the detailed chemical processes are mapped to a reduced system of tracking scalars. The governing equations of these mapped variables are much easier to handle, as the source terms do not exist any more.

\end{document}