### 2019-03-20 VG500/WEEK4/20190322.tex
weekly writting 4

\documentclass[paper=a4, fontsize=11pt]{scrartcl}
\usepackage{amsmath, amsfonts, amssymb}

\title{Paraphrase of  a paragraph in "Progress-variable approach for large-eddy simulation of non-premixed turbulent combustion"}
\author{Yu Cang \ \ 018370210001}
\date{2019-03-22}

\begin{document}

\maketitle

\section{Original paragraph}

\section{My paraphrase}

\end{document}