### 2019-04-15 VG500/HW5/submission.tex
add ex5

\documentclass[paper=a4, fontsize=11pt]{scrartcl} % A4 paper and 11pt font size

\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\usepackage{fourier} % Use the Adobe Utopia font for the document - comment this line to return to the LaTeX default
\usepackage[english]{babel} % English language/hyphenation
\usepackage{amsmath,amsfonts,amsthm,amssymb} % Math packages

\usepackage{algorithm, algorithmic}
\renewcommand{\algorithmicrequire}{\textbf{Input:}} %Use Input in the format of Algorithm  
\renewcommand{\algorithmicensure}{\textbf{Output:}} %UseOutput in the format of Algorithm

\usepackage{graphicx}
\usepackage{blindtext}
\usepackage{enumerate}
\usepackage{ulem} 
\usepackage{pdfpages}
\usepackage{multirow}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{listings}
\lstset{language=Matlab}
\usepackage{hyperref}
\usepackage{lipsum} % Used for inserting dummy 'Lorem ipsum' text into the template

\usepackage{sectsty} % Allows customizing section commands
\allsectionsfont{\centering \normalfont\scshape} % Make all sections centered, the default font and small caps

\usepackage{fancyhdr} % Custom headers and footers
\pagestyle{fancyplain} % Makes all pages in the document conform to the custom headers and footers
\fancyhead{} % No page header - if you want one, create it in the same way as the footers below
\fancyfoot[L]{} % Empty left footer
\fancyfoot[C]{} % Empty center footer
\fancyfoot[R]{\thepage} % Page numbering for right footer
\renewcommand{\headrulewidth}{0pt} % Remove header underlines
\renewcommand{\footrulewidth}{0pt} % Remove footer underlines
\setlength{\headheight}{13.6pt} % Customize the height of the header

\numberwithin{equation}{section} % Number equations within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{figure}{section} % Number figures within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{table}{section} % Number tables within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)

\setlength\parindent{0pt} % Removes all indentation from paragraphs - comment this line for an assignment with lots of text

\newcommand{\horrule}[1]{\rule{\linewidth}{#1}} % Create horizontal rule command with 1 argument of height
\newcommand*{\dif}{\mathop{}\!\mathrm{d}}

\title{	
\normalfont \normalsize 
\textsc{Shanghai Jiao Tong University, UM-SJTU JOINT INSTITUTE} \\ [25pt] % Your university, school and/or department name(s)
\horrule{0.5pt} \\[0.4cm] % Thin top horizontal rule
\huge Technical Communication\\ HW5 \\ % The assignment title
\horrule{2pt} \\[0.5cm] % Thick bottom horizontal rule
}

\author{Yu Cang \quad 018370210001} % Your name

\date{\normalsize \today} % Today's date or a custom date

\begin{document}

\maketitle % Print the title

\section{Tense}
	\begin{enumerate}
		\item \textbf{past tense}\newline
			By the time Tom noticed (notice) the doorbell, it had already rung (ring) three times. As usual, he was listening (listen) to loud music on his stereo. He turned (turn) the stereo down and stood up (stand up) to answer (answer) the door. An old man was standing (stand) on the steps. The man began (begin) to speak (speak) slowly, asking (ask) for directions.

		\item \textbf{present tense}\newline
			By the time Tom notices(notice) the doorbell, it has already rung(ring) three times. As usual, he is listening (listen) to loud music on his stereo. He turns (turn) the stereo down and stands up (stand up) to answer (answer) the door. An old man is standing (stand) on the steps. The man begins (begin) to speak (speak) slowly, asking (ask) for directions.
	\end{enumerate}

\section{Passive voice}
	\begin{enumerate}
		\item 
			Your kindness should be thankful.(no passive voice)
		\item 
			The decision has been delayed due to a lack of time.
		\item 
			Preparations for the project was begun in June.
		\item 
			Several objections to the decision were raised by the customers.
		\item 
			Your application has been processed.
	\end{enumerate}

\section{Positive tone}
	\begin{enumerate}
		\item 
			You made this mistake due to your misunderstanding of our letter.
		\item 
			It indicates that you know absolutely nothing about our ﬁeld of research from your comment on the second paragraph of page four.
		\item
			 One-tenth of the results of the experiments are incoherent.
		\item
			 Since the experimental setup was not taken enough care for, the results cannot be as precise as they could have been.
		\item
			 We reached a 99.7\% purity level, failing to achieve a level of 100\% as many other authors [5, 12, 14, 15].
	\end{enumerate}

\section{\LaTeX}
	Lorem\citep{RN5}  ipsum  dolor\citep{RN11}  sit  amet,  consectetuer  adipiscing  elit.   Ut  purus  elit,  vestibulum  ut,placerat  ac,  adipiscing  vitae,  felis.   Curabitur  dictum  gravida  mauris.   Nam  arcu  libero,nonummy eget, consectetuer id, vulputate a, magna.  Donec vehicula augue eu neque.  Pel-lentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.Mauris ut leo.  Cras viverra metus rhoncus sem.  Nulla et lectus vestibulum urna fringillaultrices.   Phasellus\citep{RN14}  eu  tellus  sit  amet  tortor  gravida  placerat.   Integer  sapien  est,  iaculisin,  pretium  quis,  viverra  ac,  nunc.   Praesent  eget  sem  vel  leo  ultrices  bibendum.   Aeneanfaucibus.  Morbi dolor nulla, malesuada eu, pulvinar at, mollis ac, nulla.  Curabitur\citep{RN10} auctorsemper  nulla.   Donec  varius  orci  eget  risus.   Duis  nibh  mi,  congue  eu,  accumsan  eleifend,sagittis quis, diam.  Duis eget orci sit amet orci dignissim rutrum.

\section{Group exercise}
	In the traditional method of two round-trip secure search, the trapdoors of each keyword list are generated by the data owner or his agent, and then the user will search them. This will be a time-consuming job for mobile device and significantly degrade the user experience. In our method, each trapdoor can be generated by its own device.\newline
	The user terminal stores the noise set and a frequently-used cache. The cache originally stores the keywords which are commonly used in the corpus and corresponding feature bits. If the user inputs a keyword, it will do the stemming process first, and then search the cache. If finding the keyword, it will return the intermediate trapdoor immediately, otherwise it will calculate the feature bits and add the keyword to the cache for next time use. The whole process is presented in Fig. 4. By introducing cache, the time of generating trapdoors can be efficiently reduced.

\bibliographystyle{plainnat}
\bibliography{reference}

\end{document}