### 2019-04-19 VG500/Slide/tex/main.tex
add intro

\input{preamble.tex}
\input{title.tex}

\begin{document}

\maketitle

\section{Introduction}

	\subsection{part1.1}

		\begin{xframe}{Backgound}
			Turbulent combustion is encountered in most practical combustion system such as \textit{rocket}, \textit{ICE}, \textit{aircraft engien}, and \textit{industrial burner}.
			\begin{figure}[htbp]\centering
				\begin{minipage}{3.5cm}
					\centering
					\includegraphics[height=4.5cm, width=3cm]{../pic/rocket.jpg}
				\end{minipage}%
				\begin{minipage}{7.5cm}
					\centering
					\includegraphics[height=4.5cm, width=7cm]{../pic/engien.jpg}
				\end{minipage}%
			\end{figure}
		\end{xframe}

		\begin{xframe}{Research aspects}
			Meaningful to practical systems:
			\begin{enumerate}[(a)]
				\item 
					Improve efficiency, meet demanding standards.
				\item
					Reduce pollution, environment friendly.
			\end{enumerate}

			Why is numerical simulation adopted?
			\begin{enumerate}[(a)]
				\item 
				Analytical techniques are difficult to handle.
				\item
				Experiments are too expensive to be widely used.
			\end{enumerate}
		\end{xframe}

		\begin{xframe}{Physical Configuration}
			\includegraphics[width=7cm, height=8cm]{../pic/counterflow_configuration.png}	
		\end{xframe}

{
\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../pic/theme-2.png}}
\begin{frame}[plain]

\end{frame}
}

{
\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../pic/theme-3.png}}
\begin{frame}[plain]

\end{frame}
}

\section{Governing Equations}

	\subsection{Showing code}

		\begin{xframe}{Code snippets}

		    Do you have some code to show on the slide?

		    And the same frame should also contain text?

		    \begin{cxxcodebox}
		        class example {
		            // \codedots shows grayed-out dots
		            @ \codedots @
		        };
		    \end{cxxcodebox}

		    You can use \verb|cxxcodebox| environment.
		    It has \verb|cxx| in the name,
		    but no syntax highlighting is performed.

		    For short code snippets,
		    it is better just to highlight the important parts.

		\end{xframe}

	\subsection{Code slides and escapes}

		\begin{xframe}{Second slide}

		    \begin{cxxcode}
		        class example {
		            // There are a few useful escapes here
		            @ \codedots @ // \codedots shows grayed-out dots

		            // Invalid parts can be marked with \hlErr
		            @\hlErr{operator;}@

		            // Good parts can be marked with \hlOk
		            @\hlOk{operator() ()}@

		            // Other highlighting commands can be seen in
		            // the preamble.tex file
		        };
		    \end{cxxcode}

		\end{xframe}

\section{Numerical Solution}

\end{document}